---
title: "Coursera_PML_Project"
author: "Ashish Jha"
date: "2 September 2017"
output: html_document
---


#Reading data into R

```{r}
library(readr)
library(caret)
library(randomForest)
training <- read_csv("pml-training.csv")
pml_testing <- read_csv("pml-testing.csv")
```

removing columns which have more than 90% missing values

```{r}
training<- training[, -which(colMeans(is.na(training)) > 0.9)]
pml_testing<- pml_testing[, -which(colMeans(is.na(pml_testing)) > 0.9)]
```

Coverting classe in dataset as dummy variables because classe is of character dataset and would pose difficulty in further analysis.

```{r}
dummy <- training$classe
dummy[dummy == "A"] <- 1
dummy[dummy == "B"] <- 2
dummy[dummy == "C"] <- 3
dummy[dummy == "D"] <- 4
dummy[dummy == "E"] <- 5

dummy <- as.numeric(dummy)

training$dummy <- dummy
```

Spliting dataset into training set and validation set

```{r}
part <- createDataPartition(y=training$classe, p=0.7, list = FALSE)
train <- training[part,]
valid <- training[-part,]
```



Checking datatype of varibles in dataset since variables with character datatype won't work with randomforest
```{r}
str(train)
```



Using random forest to train the using randomforest with 500 trees. Used randomforest as it is work horse , tried and tested. RandomForest works well all the time.


```{r}
fit_train <- randomForest(formula = dummy ~ ., data = train[,-c(1,2,5,6,60)],importance=TRUE,na.action=na.omit,ntree = 500,mtry=100)
```

#Prediciting on validation set

```{r}
pred_valid <- predict(fit_train, newdata = valid[,-c(1,2,5,6,60)])
pred_valid_round <- round(pred_valid, digits = 0)
table(pred_valid_round, valid$dummy)
```

Test results were overwhelming with accuracy of 99.67% . Hence i will accept this model for final prediction 

```{r}
pred_test <- predict(fit_train, newdata= pml_testing[,-c(1,2,5,6,60)])
pred_test_round <- round(pred_test, digits = 0)
pred_test_round

```



















